---
title: "ESM 232 Assignment 7"
author: "Madeline Oliver, Alex Ehrens, Siya Qiu"
date: "5/12/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sensitivity)
```

# Rabbits and Hawks Population Matrix Assignment


```{r}

source("evolve_pop2.R")

# fertility rates
f1 = 0
f2 = 2/2 # have to divide by two because only women are fertile
f3 = 6/2
f4 = 1/2

# survivability 
p01 = 0.8
p12 = 0.85
p23 = 0.65
p34 = 0.1

# initial population parameters
ini = c(0,0,10,0) 
nyears = 20
fert_rabbits = c(f1,f2,f3,f4)
surv_rabbits = c(p01,p12,p23,p34)
rabbit_pop=evolve_pop2(fert_rabbits, surv_rabbits, ini, nyears)

head(rabbit_pop)

```
 
### Matrix population model result:

After 20 years and an initial population of 10 adult (age 2-3) rabbits, there are about 36,452 rabbits in the total population, with about 18,372 rabbits in the young (0-1) age class.

### Sobel sensitivity analysis

We use Sobel to explore how the intervention of encouraging nesting of hawks that eat rabbits might impact the rabbit population. We also account for uncertainty in our survivability parameters.

```{r sobel}
# survivability 
nsample=200

# create our two samples for Sobel
# first do our survivability
ps1 = cbind.data.frame(p01 = runif(min=0.65, max=0.75, n=nsample), 
                       p12 = runif(min=0.75, max=0.8, n=nsample),
                       p23 = runif(min=0.65, max=0.65, n=nsample),
                       p34 = runif(min=0.1, max=0.1, n=nsample))

ps2 = cbind.data.frame(p01=runif(min=0.65, max=0.75, n=nsample), 
                       p12 = runif(min=0.75, max=0.8, n=nsample),
                       p23 = runif(min=0.65, max=0.65, n=nsample),
                       p34 = runif(min=0.1, max=0.1, n=nsample))

# then fertility
fs1 = cbind.data.frame(f1 = 0, 
                       f2 = 1,
                       f3 = 3,
                       f4 = 0.5)

# put survivability and fertility together
allp1 = cbind.data.frame(ps1,fs1)
allp2 = cbind.data.frame(ps2,fs1)

# get sobel samples
sens_rabbit=soboljansen(model = NULL, allp1, allp2, nboot = 100)

head(sens_rabbit$X)
nsim=nrow(sens_rabbit$X)

# run model and save what we care about: final population after 20 years
# this is already output by evolve_pop so we don't need a compute_metric function

ini = c(0,0,10,0)
nyears=20

# as before combine our application of the the dynamics model - for each
# parameter set, with code to extract our metric of interest (final population)
p_wrapper = function(p01, p12, p23, p34, f1,f2,f3,f4,use_func, initialpop, nstep ) {
  fertility=c(f1,f2,f3,f4)
  survivability= c(p01,p12,p23,p34)
  res = use_func(survivability =survivability, fertility = fertility, initialpop=initialpop, nstep=nstep)
# now return the final population total
  return(finalpop=res$poptot[nstep])
}

# use pmap here so we can specify rows of our sensitivity analysis parameter object 
res = as.data.frame(sens_rabbit$X) %>% 
  pmap_dbl(p_wrapper, initialpop=ini, nstep=nyears, use_func=evolve_pop2)

# plot results (variation in final population across all parameter)
# ggplot needs a dataframe - so do a quick conversion with data.frame
ggplot(data.frame(finalpop=res), 
       aes(x=finalpop))+
  geom_density()

# boxplot of variation in total rabbit population
ggplot(data.frame(finalpop=res), 
       aes(x="", y=finalpop) )+
  geom_boxplot(fill="blue")+
  theme(axis.title.x = element_blank())+
  labs(y="Final Population")

# give our results to sensitivity structure
sens_rabbit=tell(sens_rabbit, res)

# look at results
sens_rabbit$S
sens_rabbit$T

# Graph how total rabbit population after 20 years varies with survivability of young age and sub-adult
tmp = cbind.data.frame(sens_rabbit$X, 
                       pop12=sens_rabbit$y)

# plot of total rabbit population vs. survivability of young
ggplot(tmp, aes(p01, pop12))+
  geom_point()+
  labs(x="Survivability of young",
       y="Population after 20 years")

# plot of total rabbit population vs. survivability of sub-adult
ggplot(tmp, aes(p12, pop12))+
  geom_point()+
  labs(x="Survivability of sub-adults",
       y="Population after 20 years")

# total population after 20 years
#total populations at each sensitivity level

head(res)

#p01
tmp_p01 <- tmp %>%
  mutate(p01 = round(p01, digits = 2)) %>% 
  group_by(p01) %>% 
  summarise(pop = mean(pop12))
#p12
tmp_p12 <- tmp %>%
  mutate(p12 = round(p12, digits = 2)) %>% 
  group_by(p12) %>% 
  summarise(pop = mean(pop12))

```

When we include the hawk intervention in our population model and account for uncertainty in survivability rates, our population after 20 years 
